\chapter{Other: this needs a better title}

\section{Presenting Your Reactor to Others}
There comes a dreaded point in time where a reactor design must be communicated to others.
For the introverted engineer this can be a painful process.
For the student presenting a capstone project as a final presentation or at a conference the stakes may appear quite high.
Thankfully, such presentations do not need to be so feared. Proper preparation can help.

Presenting a reactor design is difficult to do in the short amount of time allotted for senior design presentations and conference paper presentations. 
Nuclear reactors are complex, multi-faceted systems which cannot be fully described in 20 minutes. This makes them an easy target for tough questions, which usually begin with ``Did you consider...'' 
Clearly it is impossible to consider everything beforehand unless this is a very mature design. 
Often it is perfectly acceptable to relegate such questions to ``future work'' or ``beyond the scope of this presentation,'' but cogent answers will demonstrate your understanding of nuclear reactor design. Even if you haven't thought of that particular effect, the audience would like to see you give a thoughtful reply about how it might influence your future design.

Whatever you do, don't become defensive.

Preparation is important. With good preparation and practice, it will be much easier to be confident. A confident presenter appears much more competent and also makes the audience much more at ease. An air of confidence also makes responses to questions sound much more final, so that questioners do not continue piling on questions in the same vein.
Good preparation will also include asking for feedback from a friendly audience. 
Have them expose the obvious flaws in your logic and presentation.
While you cannot consider everything, make sure you have considered everything that could ruin your design. ``Have you considered the price of unobtainium in your economics analysis?''
The capabilities that unobtainium provides must merit its use in your reactor. 
``Will the adamantium corrode in the presence of vaporized sodium coolant?'' The world doesn't need sci-fi reactors. Do the research necessary to show your audience that you have a serious design. 

Remember that no reactor is perfect. Explain the weaknesses of your design honestly.
Then you can argue that the capabilities and performance of your design merit the R\&D effort to overcome these problems.
